---
title: "Pain in Patients with Non-small Cell Lung Cancer"
format: docx
author: 
  - name: Shelly Li
execute:
  echo: false
  warning: false
  error: false
date: "`r format(Sys.Date(), '%Y%m%d')`"
output-file: "ML-DataReport"
---

```{r}
#| echo: false
library(tidyverse)
library(gt)
library(gtsummary)
library(scales)
library(texreg)
library(psych)
library(knitr)
load("/Volumes/jim_gonzalez_research/DATA/Jim/Machine Learning/Data Analysis/Li - Pain/AnalysisData.RData")
```

Data were collected from 162 patients between `r min(scored$dateofcompletion, na.rm = TRUE)` and `r max(scored$dateofcompletion, na.rm = TRUE)`. Based on the protocol, we should have collected baseline data before or at the date of the first immunotherapy. However, for some patients, baseline dates happened after the first immunotherapy dates. 

-  Here is the frequency table for the number of days from the first immunotherapy (IMMUNO1_DTE) to baseline assessment (baselinedate). Note that bl_tx1 = IMMUNO1_DTE - baselinedate, and those with negative days indicate baseline data collected after the first treatment (n = 12).

```{r}
#|label: tbl-bltotx
test %>%
  mutate(bl_tx1_factor = as.factor(bl_tx1)) %>%
  tbl_summary(by = NULL, 
              include = bl_tx1_factor, 
              statistic = list(all_categorical() ~ "{n} ({p}%)"))
```

## The following 22 patients look problematic

They either:

1.  appeared to have baseline after the date of first treatment (n = 12, e.g., studyid = 1), or

2.  were missing date of baseline or date of first treatment (n = 7, e,g., studyid = 8), or

3.  have the first treatment more than 3 months after they took baseline (n = 3, studyid = 47, 48, or 51).

### Can coordinators look into these cases？

```{r}
#|label: tbl-problem

list<-test %>%
  select(-date_immuno1) %>%
  mutate(baselinedate = format(baselinedate, "%m/%d/%Y")) %>%
  filter(bl_tx1 < 0 | bl_tx1 > 21 | is.na(bl_tx1)) %>%
  as.data.frame()

print(list)
```

-  We also have 10 patients whose baseline assessments were collected between 7 to 21 days before the first treatment. These should not be considered as “baseline” pain assessments.

```{r}
#|label: tbl-problem

list<-test %>%
  select(-date_immuno1) %>%
  mutate(baselinedate = format(baselinedate, "%m/%d/%Y")) %>%
  filter(bl_tx1 %in% (7:21)) %>%
  as.data.frame()

print(list)
```
